### 2017-07-17 slowtests/slowtests-boston.Rmd
switching to knitr

---
title: "Bouston Housing Tests"
author: "Brandon M. Greenwell"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
# Load required packages
library(ggplot2)
library(grid)
library(pdp)

# Set knitr chunk options
knitr::opts_chunk$set(echo = TRUE)

# Function to run tests for a fitted model object using the Boston housing data
bostonHousingTest <- function(object, ...) {

  # Training data
  trn <- subset(boston, select = -cmedv)

  # Partial dependence plot for lstat
  pd1 <- partial(object, pred.var = "lstat", train = trn, ...)

  # Partial dependence plot for lstat and rm
  pd2 <- partial(object, pred.var = c("lstat", "rm"), chull = TRUE,
                 train = trn, ...)

  # Individual conditional expectation curves for age
  pd3 <- partial(object, pred.var = "age", ice = TRUE, train = trn, ...)

  # Centered individual conditional expectation curves for lstat
  pd4 <- partial(object, pred.var = "age", ice = TRUE, center = TRUE,
                 train = trn, ...)

  # Display plots
  grid.arrange(
    autoplot(pd1, rug = TRUE, train = trn, main = "PDP for lstat"),
    autoplot(pd2, main = "PDP for lstat and rm"),
    autoplot(pd3, main = "ICE curves for age", alpha = 0.25),
    autoplot(pd4, main = "c-ICE curves for age", alpha = 0.25),
    ncol = 2, 
    top = textGrob(deparse(substitute(object)),
                   gp = gpar(fontsize = 20, font = 3))
  )

}
```

## Cubist

# `Cubist` package

```{r}
boston.cubist <- Cubist::cubist(x = subset(boston, select = -cmedv),
                                y = boston$cmedv, committees = 100)
bostonHousingTest(boston.cubist)
```

## Decision trees

# `rpart` package

```{r}
boston.rpart <- rpart::rpart(cmedv ~ ., data = boston)
bostonHousingTest(boston.rpart)
```

# `party` package

```{r}
boston.ctree <- party::ctree(cmedv ~ ., data = boston)
bostonHousingTest(boston.ctree)
```

# `partykit` package

```{r}
boston.ctree2 <- partykit::ctree(cmedv ~ ., data = boston)
bostonHousingTest(boston.ctree2)
```

## Bagging

# `ipred` package

```{r}
set.seed(101)
boston.ipred <- ipred::bagging(cmedv ~ ., data = boston, nbagg = 500)
bostonHousingTest(boston.ipred)
```

## Random forests

# `randomForest` package

```{r}
set.seed(101)
boston.rf <- randomForest::randomForest(cmedv ~ ., data = boston)
bostonHousingTest(boston.rf)
```

# `party` package

```{r}
set.seed(101)
boston.crf <- party::cforest(cmedv ~ ., data = boston)
bostonHousingTest(boston.crf, quantiles = TRUE)
```

# `ranger` package

```{r}
set.seed(101)
boston.ranger <- ranger::ranger(cmedv ~ ., data = boston)
bostonHousingTest(boston.ranger)
```

## Boosting

# `gbm` package

```{r}
set.seed(101)
boston.gbm <- gbm::gbm(cmedv ~ ., data = boston, distribution = "gaussian",
                       n.trees = 5000, interaction.depth = 3, shrinkage = 0.001,
                       cv.folds = 5)
best.iter <- gbm::gbm.perf(boston.gbm, method = "OOB", plot.it = FALSE)
bostonHousingTest(boston.gbm, n.trees = best.iter)
```

# `xgboost` package
```{r}
set.seed(101)
boston.xgb <- xgboost::xgboost(
  data = data.matrix(subset(boston, select = -cmedv)), label = boston$cmedv,
  objective = "reg:linear", nrounds = 5000, max_depth = 3, eta = 0.001,
  colsample_bytree = 0.8, subsample = 0.5,
  save_period = NULL, verbose = 0
)
bostonHousingTest(boston.xgb)
```

## Neural networks

# `nnet` package
```{r}
set.seed(101)
boston.nnet <- nnet::nnet(cmedv ~ ., data = boston, size = 6, decay = 0.1,
                          linout = TRUE, trace = FALSE)
bostonHousingTest(boston.nnet)
```

## Support vector machines

# `e1071` package

```{r}
boston.svm <- e1071::svm(cmedv ~ ., data = boston, type = "eps-regression")
bostonHousingTest(boston.svm)
```

# `kernlab` package

```{r}
boston.ksvm <- kernlab::ksvm(cmedv ~ ., data = boston, type= "eps-svr")
bostonHousingTest(boston.ksvm)
```

## Linear/generalized linear models

# `stats` package

```{r}
boston.lm <- lm(cmedv ~ . ^ 2, data = boston)
bostonHousingTest(boston.lm)
```

```{r}
boston.glm <- glm(cmedv ~ . ^ 2, data = boston)
bostonHousingTest(boston.glm)
```

## Multivariate adaptive regression splines

# `earth` package

```{r}
boston.earth <- earth::earth(cmedv ~ ., degree = 3, data = boston)
bostonHousingTest(boston.earth)
```

# `mda` package

```{r}
boston.mars <- mda::mars(x = subset(boston, select = - cmedv), 
                         y = boston$cmedv, degree = 3)
bostonHousingTest(boston.mars)
```