### 2017-07-18 slowtests/slowtests-pima.Rmd
slowtests (mostly) updated

---
title: "Bouston  Tests"
author: "Brandon M. Greenwell"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
# Load required packages
library(ggplot2)
library(grid)
library(pdp)

# Set knitr chunk options
knitr::opts_chunk$set(echo = TRUE)

# Function to run tests for a fitted model object using the Pima Indians 
# diabetes data
pimaIndiansTest <- function(object, ...) {

  # Training data
  pima <- na.omit(pima)
  trn <- subset(pima, select = -diabetes)

  # Partial dependence plot for lstat
  pd1 <- partial(object, pred.var = "glucose", train = trn, ...)

  # Partial dependence plot for lstat and rm
  pd2 <- partial(object, pred.var = c("glucose", "age"), chull = TRUE,
                 train = trn, ...)

  # Individual conditional expectation curves for age
  pd3 <- partial(object, pred.var = "glucose", ice = TRUE, train = trn,
                 ...)

  # Centered individual conditional expectation curves for lstat
  pd4 <- partial(object, pred.var = "glucose", ice = TRUE, center = TRUE,
                 train = trn, ...)

  # Display plots
  grid.arrange(
    autoplot(pd1, rug = TRUE, train = pima, main = "PDP for lstat"),
    autoplot(pd2, main = "PDP for glucose and age"),
    autoplot(pd3, rug = TRUE, train = pima,
             main = "ICE curves for glucose", alpha = 0.25),
    autoplot(pd4, rug = TRUE, train = pima,
             main = "c-ICE curves for glucose", alpha = 0.25),
    ncol = 2,
    top = textGrob(deparse(substitute(object)),
                   gp = gpar(fontsize = 20, font = 3))
  )

}
```

# Discriminant analysis

## Package: `MASS`

```{r}
pima.lda <- MASS::lda(diabetes ~ . ^ 2, data = na.omit(pima))
pimaIndiansTest(pima.lda)
pimaIndiansTest(pima.lda, prob = TRUE)
```

```{r}
pima.qda <- MASS::qda(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.qda)
pimaIndiansTest(pima.qda, prob = TRUE)
```

## Package: `mda`

```{r}
pima.mda <- mda::mda(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.mda)
pimaIndiansTest(pima.mda, prob = TRUE)
```

```{r}
pima.fda <- mda::fda(diabetes ~ ., data = na.omit(pima), method = mda::mars,
                     degree = 2)
pimaIndiansTest(pima.fda)
pimaIndiansTest(pima.fda, prob = TRUE)
```

# Decision trees

## Package: `rpart`

```{r}
pima.rpart <- rpart::rpart(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.rpart)
pimaIndiansTest(pima.rpart, prob = TRUE)
```

## Package: `C50`
```{r}
set.seed(101)
pima.C5.0 <- C50::C5.0(diabetes ~ ., data = na.omit(pima), trials = 100)
pimaIndiansTest(pima.C5.0)
pimaIndiansTest(pima.C5.0, prob = TRUE)
```

## Package: `party

```{r}
pima.ctree <- party::ctree(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.ctree)
pimaIndiansTest(pima.ctree, prob = TRUE)
```

## Package: `partykit

```{r}
pima.ctree2 <- partykit::ctree(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.ctree2)
pimaIndiansTest(pima.ctree2, prob = TRUE)
```

# Bagging

## Package: `adabag`

```{r}
set.seed(101)
pima.bagging <- adabag::bagging(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.bagging, quantiles = TRUE)
pimaIndiansTest(pima.bagging, quantiles = TRUE, prob = TRUE)
```

## Package: `ipred

```{r}
set.seed(101)
pima.ipred <- ipred::bagging(diabetes ~ ., data = na.omit(pima), nbagg = 500)
pimaIndiansTest(pima.ipred, quantiles = TRUE)
pimaIndiansTest(pima.ipred, quantiles = TRUE, prob = TRUE)
```

# Random forests

## Package: `randomForest

```{r}
set.seed(101)
pima.rf <- randomForest::randomForest(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.rf)
pimaIndiansTest(pima.rf, prob = TRUE)
```

## Package: `party

```{r}
set.seed(101)
pima.crf <- party::cforest(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.crf, quantiles = TRUE)
pimaIndiansTest(pima.crf, quantiles = TRUE, prob = TRUE)
```

## Package: `partykit`
```{r}
set.seed(101)
pima.crf2 <- partykit::cforest(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.crf2, quantiles = TRUE)
pimaIndiansTest(pima.crf2, quantiles = TRUE, prob = TRUE)
```

## Package: `ranger`

```{r}
set.seed(101)
pima.ranger <- ranger::ranger(diabetes ~ ., data = na.omit(pima), 
                              probability = TRUE)
pimaIndiansTest(pima.ranger)
pimaIndiansTest(pima.ranger, prob = TRUE)
```

# Boosting

## Package: `adabag`

```{r}
set.seed(101)
pima.boosting <- adabag::boosting(diabetes ~ ., data = na.omit(pima))
pimaIndiansTest(pima.boosting, quantiles = TRUE)
pimaIndiansTest(pima.boosting, quantiles = TRUE, prob = TRUE)
```

## Package: `gbm`

```{r}
set.seed(101)
pima.gbm <- gbm::gbm(diabetes ~ ., 
                     data = na.omit(pima),
                     distribution = "multinomial",
                     n.trees = 5000,
                     interaction.depth = 3,
                     shrinkage = 0.001,
                     cv.folds = 5,
                     verbose = FALSE)
best.iter <- gbm::gbm.perf(pima.gbm, method = "cv", plot.it = FALSE)
pimaIndiansTest(pima.gbm, n.trees = best.iter)
pimaIndiansTest(pima.gbm, n.trees = best.iter, prob = TRUE)
```

## Package: `xgboost`
```{r}
set.seed(101)
pima.xgb <- xgboost::xgboost(
  data = data.matrix(subset(pima, select = -diabetes)),
  label = unclass(pima$diabetes) - 1, objective = "binary:logistic",
  nrounds = 100, max_depth = 3, eta = 0.1, gamma = 0, colsample_bytree = 0.8, 
  min_child_weight = 1, subsample = 0.7, verbose = 0
)
pimaIndiansTest(pima.xgb)
pimaIndiansTest(pima.xgb, prob = TRUE)
```

# Neural networks

## Package: `nnet`
```{r}
set.seed(101)
pima.nnet <- nnet::nnet(diabetes ~ ., data = na.omit(pima), size = 10, 
                        decay = 0.1, maxit = 500, trace = FALSE)
pimaIndiansTest(pima.nnet)
pimaIndiansTest(pima.nnet, prob = TRUE)
```

# Support vector machines

## Package: `e1071`

```{r}
pima.svm <- e1071::svm(diabetes ~ ., data = na.omit(pima), type = "C-classification", 
                       probability = TRUE)
pimaIndiansTest(pima.svm)
pimaIndiansTest(pima.svm, prob = TRUE)
```

## Package: `kernlab`

```{r}
pima.ksvm <- kernlab::ksvm(diabetes ~ ., data = na.omit(pima), type = "C-svc", 
                           prob.model = TRUE)
pimaIndiansTest(pima.ksvm)
pimaIndiansTest(pima.ksvm, prob = TRUE)
```

# Multinomial models

## Package: `nnet`

```{r}
set.seed(101)
pima.multinom <- nnet::multinom(diabetes ~ . ^ 2, data = na.omit(pima), 
                                trace = FALSE)
pimaIndiansTest(pima.multinom)
pimaIndiansTest(pima.multinom, prob = TRUE)
```